---
title: "Math 185 Homework 5"
author: "Lucien Chen"
date: "2024-05-25"
output: html_document
---

## Homework 5

``` {r setup, echo=TRUE}
library(readr)
library(dplyr)
library(ggplot2)
df <- read.csv("cars.csv") %>% select(-c(Engine.Information.Engine.Type, Identification.ID, Identification.Model.Year))
```

### Question 1

``` {r test, echo=TRUE}
cols <- names(df %>% select(-Fuel.Information.City.mpg))
response <- df$Fuel.Information.City.mpg

p_vals <- list()

for (col in cols) {
  predictor <- df[[col]]
  if (is.numeric(predictor)) {
    test <- cor.test(response, predictor)
    p_vals[col] <- test$p.value
  } else if (is.factor(predictor)) {
    test <- aov(response ~ predictor)
    p_vals[col] <- test$p.value
  }
}

p_values <- data.frame(
  predictor = names(p_vals),
  p.value = unlist(p_vals)
)

p_values
```

``` {r correction, echo=TRUE}
p_values$FWER_correction <- p.adjust(p_values$p.value, method="bonferroni")
p_values$FDR_correction <- p.adjust(p_values$p.value, method="BH")
ggplot(p_values, aes(x=predictor, y=FWER_correction)) + geom_bar(stat="identity") + geom_text(aes(label = FWER_correction), vjust = -0.5) + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="p-values with FWER Correction")
ggplot(p_values, aes(x=predictor, y=FDR_correction)) + geom_bar(stat="identity") + geom_text(aes(label = FDR_correction), vjust = -0.5)+ theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title="p-values with FDR Correction")
```

As we can see from the results of our corrections, the Bonferroni procedure is much more conservative than the Benjamini-Hochberg procedure. The p-values for the variables are much higher, i.e Dimensions.Length and Engine.Information.Number.of.Forward.Gears, and so at a given alpha level, say $\alpha=0.05$, we would end up failing to reject number of forward gears as a significant predictor, whereas under the BH procedure, we would reject. In both cases, we would fail to reject length as a significant predictor although in the Bonferroni procedure, the p-value becomes much higher (close to 1).

### Question 2

```{r association, echo=TRUE}
response <- df$Fuel.Information.City.mpg
predictor <- df$Fuel.Information.Highway.mpg

corr <- cor(response, predictor)
p_value <- cor.test(response, predictor)$p.value
print(c(corr, p_value))

ggplot(df, aes(x=Fuel.Information.Highway.mpg, y=Fuel.Information.City.mpg, color="red")) + geom_point() + theme_minimal() + labs(title="Highway vs City MPG")

lr <- lm(response ~ predictor)
summary(lr)

ggplot(df, aes(x=Fuel.Information.Highway.mpg, y=Fuel.Information.City.mpg)) + geom_point() + geom_smooth(method="lm", se=F, color="red") + labs(title="Fit of Linear Model Against the Data") + theme_minimal()
```

As we can see from the above code, the association is quite high, with a correlation of ~0.866 between the highway and city mpg. We can also see that our cor test has a p-value of 0 which indicates that there is significant association between the two variables. In the scatterplot, we can see that there is quite a good linear relationship between the variables with the exception of one outlier on the x-axis. Our linear regression also fits the data quite well, barring the outlier mentioned before (however this is likely to be an error).

### Question 3

``` {r function, echo=TRUE}
association_analysis <- function(data) {
  cols <- names(data)
  x <- data[[cols[1]]]
  y <- data[[cols[2]]]
  corr <- cor(x, y)
  p_value <- cor.test(x, y)$p.value
  print(c(corr, p_value))
  scatter <- ggplot(mapping=aes(x=x, y=y, color="red")) + geom_point() + theme_minimal() + labs(title=paste(cols[1], " vs ", cols[2]))
  print(scatter)
  lr <- lm(y ~ x)
  print(summary(lr))
  lr_plot <- ggplot(mapping=aes(x=x, y=y)) + geom_point() +  geom_smooth(method="lm", se=F, color="red") + theme_minimal() + labs(title=paste("Line of best fit for", cols[1], "and", cols[2]))
  print(lr_plot)
  return()
}

paired_data <- df %>% select(c(Fuel.Information.Highway.mpg, Fuel.Information.City.mpg))
association_analysis(paired_data)
```

As we can see, our function produces the same output as in Question 2.